# Getting started

## Obtaining and saving API Key

CMAP hosts data on a Microsoft SQL Server. To retrieve data from the database,
`cmap4r` requires an API key. For this, the user needs to register on the
[SimonsCMAP](https://simonscmap.com/register) website and navigate to the
"APIKeys" tab on the webpage to obtain an API key. The package allows the user
to save the key as a **keyring variable** which will be used for data retrieval
whenever required.

## Installation and setup

First, open R in a command line or in an R console. In order to install the R
package, first install the
[devtools](https://cran.r-project.org/web/packages/devtools/readme/README.html)
R package using `install.packages("devtools")`.

Next, install the [cmap4r](https://github.com/simonscmap/cmap4r) R package using
`devtools::install_github("simonscmap/cmap4r/cmap4r")`. There are some other
dependencies to install as well:

```{r, eval=FALSE, echo=TRUE}
## Package for data processing:
install.packages("dplyr")  

## Packages for visualization:
install.packages("ggplot2")
install.packages("plotly")
```

Now, in your R session, run `set_authorization()` to set/reset the API Key. You
will be prompted to enter the key on the command line or in a pop-up
prompt. Alternatively, you can attempt to use any of the functions
(e.g. `get_catalog()`) and you will be prompted to enter the key.

## Catalog of data

You can obtain the latest version of the data catalog for the CMAP data
(https://cmap.readthedocs.io/en/latest/catalog/catalog.html) to R using
`get_catalog()`, partially reproduced here.

```{r example-catalog, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE}
get_catalog() %>%
  select(Variable, Table_Name, Unit, Sensor, Unit) %>%
  head(20) %>%
  kable()
```
