```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8,fig.align = "center",
                      out.width = "100%")
library(DBI)
library(dplyr)
apiKey = "5e05c500-d68d-11e9-9d3b-4f83fcec4710"
## library(cmap4r)
```
# Data Retrieval

## Obtaining data summaries

Data in the CMAP database exists in tables, and can be accessed by specific
table names, and variable names therein. For instance, sea surface temperature
is from the table `"tblSST_AVHRR_OI_NRT"`, and the variable name is `"sst"`.

Some simpler functions for accessing data are useful for obtaining summaries of
the data prior to downloading the data itself.

+ Name of variables (`get_columns()`)
+ Type of variables, i.e., quantitative, qualitative, time, (`get_var_catalog()`)
+ Space-time range information (`get_var_coverage`),
+ Numerical variable range summary (`get_var_stat()`).

These are used in the various following functions, which are designed to
summarize data tables by extracting these table attributes. This is useful for
learning about the tables without downloading them prior to analysis. Here are
some key examples (the piping using `%>%` to `kable()` is just for formatting
purposes):


```{r simpler-functions, eval=TRUE, echo=TRUE}
get_head("tblArgoMerge_REP") %>% kable() 
get_var_coverage("tblAltimetry_REP", "sla") %>% kable()
get_var_resolution("tblAltimetry_REP", "sla") %>% kable()
get_var_unit("tblAltimetry_REP", "sla") %>% kable()
get_columns('tblSST_AVHRR_OI_NRT') %>% kable()
```

## Cruises

Out of the many data tables in CMAP, some are data collected and processed from
cruises. Here are some examples of summarizing this data.

```{r simpler-functions-cruises, eval=TRUE, echo=TRUE}
cruisename = "KOK1606" 
get_cruise_by_name(cruisename) %>% kable()
get_cruise_bounds(cruisename) %>% kable()
get_head('tblAMT13_Chisholm') %>% kable()
get_columns('tblAMT13_Chisholm') %>% head(10) %>% kable()
```

As you can see, each table contains many data "rows", which are indexed by
"keys" of the four values:

**(Time, Latitude, Longitude, Depth)**

Most functions in this package will involve using these data key values, or
user-specified ranges of these keys, for certain operations. The most basic
operation, explained next, is to download a dataset from a time and space range
of interest.

## Download data

When retrieving all data from **CMAP** in a "rectangle" of space and time,
specifies the following:

+ Name of table (`table`)
+ Names of variables in that table (`variable`)
+ Upper and lower limits of the variables (`dt1`, `dt2`, `lon1`, `lon2`, `lat1`,
  `lat2`, `depth1`, `depth2`)

These are used in the `space_time()` function, which downloads the data as a data
frame.

```{r spacetime, message=FALSE, warning=FALSE, echo=TRUE}
get_spacetime(tableName = 'tblArgoMerge_REP',
              varName = 'argo_merge_salinity_adj',
              dt1='2015-05-01',
              dt2='2015-05-30',
              lat1=28,
              lat2=38,
              lon1=-71,
              lon2=-50,
              depth1=0,
              depth2=100) %>% head(10) %>% kable()
```


## Colocalize data

First of all, what does it mean to *colocalize*? Imagine a ship moving along a
trajectory, whose location in space and time is recorded in
latitude/longitude/time. Then, we'd like to obtain data like sea surface
temperature or salinity, *as if we were recording it along the ship's
trajectory*. Colocalization simply means retrieving such data manually from the
database, by querying from the *vicinity* (a rectangle) of each lat/lon/time
triplet of interest, and summarizing it (e.g. taking the average).

The size of the vicinity is define by the user-specified "slack" in longitude,
latitude and time. For instance, one could ask for 1 degrees of slack in
longitude and latitude, and 6 hours of slack in time.

The main function is `along_track`, which colocalizes data for one or multiple
tables and variables at a time. The available cruises can be found by running
the function `cruises()`. The available data tables and variable names are
listed in the catalog (`get_catalog()`, or from
https://cmap.readthedocs.io/en/latest/catalog/catalog.html). Here is a short
example:

```{r colocalize, message=FALSE, eval=TRUE, echo=TRUE}
cruise = 'diel'
targetTables = c('tblSeaFlow', 'tblPisces_NRT')
targetVars = c('synecho_abundance', 'NO3')
depth1 = 0
depth2 = 5
temporalTolerance = c(0, 4)
latTolerance = c(0, 0.25)
lonTolerance = c(0, 0.25)
depthTolerance = c(5, 5)
dat = along_track(cruise[1],
                  targetTables[1],
                  targetVars[1],
                  depth1[1],
                  depth2[1],
                  temporalTolerance[1],
                  latTolerance[1],
                  lonTolerance[1],
                  depthTolerance[1])
```
```{r colocalize-print, echo=TRUE, eval=TRUE}
head(dat) %>% kable()
```

## Issue SQL query

(For advanced users) Simons CMAP datasets are hosted in a SQL database and the
CMAP4R R package provides the user with a number of pre-developed methods to
extract and retrieve subsets of the data. The rest of this documentation is
dedicated to explore and explain these methods. In addition to the pre-developed
functions described above, you may issue custom SQL query statement and returns
the results in form of a data frame.


```{r manual-query, eval=TRUE}
dat = exec_manualquery("SELECT [time], lat, lon, depth, Fe FROM tblPisces_NRT
         WHERE
         [time] BETWEEN '2017-06-03' AND '2017-06-03' AND
         lat BETWEEN 10 AND 55 AND
         lon BETWEEN -180 AND 100 AND
         depth BETWEEN 0 AND 0.5
         ORDER BY [time], lat, lon, depth") 
dat %>% head(10) %>% kable()
```
