# Visualization

There are two built-in options for making plots -- scatter plots to compare two
variable of interest, or histograms of individual variables.

## Scatterplot of two variables

**Code is outdated** 

The `plot_xy()` function makes a scatterplot of two variables (from two tables),
aggregated across a variable called `agg_var`. Why aggregate though? The two
variables from the two tables might exist at different time/space
resolutions. Because of this potential discrepancy in data resolution, averaging
the two datasets at some resolution according to, say, `agg_var="time"`, by day,
is a useful step before plotting a relationship between the two.

There are currently two methods for plotting -- one using the `ggplot2` R
package, and another using `plotly`.

<!-- **plot_xy** function download the data, and output a list which contain a) plot -->
<!-- object: **plot_ly/ggplot**; b) corresponding data tables. -->

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=FALSE}
con <- dbConnect(odbc::odbc(), DSN="CMAP-MSSQL",UID="ArmLab",PWD="ArmLab2018")

## Define variables
table_list <- c('tblSST_AVHRR_OI_NRT', 'tblAltimetry_REP') 
var_list <-  c('sst', 'sla')  
range_var <- list()
range_var$lat <- c(25,30)
range_var$lon <- c(-160, -155)
range_var$time <- c('2016-03-29', '2016-05-29')
agg_var <- 'time' ## agg_var <- 'lat'

## -----------------------------------
## Dataset from table I
selIndex <- 1
table_name <- table_list[selIndex]                       # Specify table name I
sel_var <- var_list[selIndex]                            # Variable from table name I  
tbl_subset_x <- get_aggtable(con, table_name, sel_var, range_var, agg_var)
head(tbl_subset_x)

## Dataset from table II
selIndex <- 2
table_name <- table_list[selIndex]                       # Specify table name II
sel_var <- var_list[selIndex]                            # Variable from table name II
tbl_subset_y <- get_aggtable(con, table_name, sel_var, range_var, agg_var)
head(tbl_subset_y)

## Plot
out <- plot_xy(con, table_list, var_list, range_var, agg_var, type = 'plotly')
out$plot
dbDisconnect(con)
```

## Histograms 

**Code is outdated** 

To generate a histogram plot of a single variable of interest, simply specify which variable and which time/lat/lon/depth range, and use the function `plot_hist()` as follows:

```{r message=FALSE, warning=FALSE, echo=TRUE, eval=FALSE}
con <- dbConnect(odbc::odbc(), DSN="CMAP-MSSQL",UID="ArmLab",PWD="ArmLab2018")

## Specify what to query
table_name <- c('tblSST_AVHRR_OI_NRT')
sel_name <-  c('sst')
range_var <- list()
range_var$lat <- c(20, 24)
range_var$lon <- c(-170, 150)
range_var$time <- c('2016-04-30', '2016-04-30')

## Obtain a data table:
tbl_subset <- get_table(con, table_name, sel_var, range_var)
head(tbl_subset)

## Make plot
p <- plot_hist(tbl_subset, 'plotly', sel_var)
p

dbDisconnect(con)
```
