# Examples

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=TRUE, eval=TRUE, cache=TRUE,
                      warning=FALSE, message=FALSE)
library(ggplot2)
library(tidyverse)
library(arrow)
library(lubridate)
library(dplyr)
library(cmap4r)
```

## Compare in-situ (on-site) sunlight data to remote sensing (satellite) data.

The intensity of sunlight is usually measured as "Photosynetheically Active
Radiation", or PAR for short.

First, we query PAR from two sources of data -- one *colocalized* from a remote
sensing (satellite) table, and one from an in-situ (on-site) cruise data table
in Simons CMAP.

The colocalization is done as follows:

```{r query1}
tabname = "tblModis_PAR"
varname = "PAR"
<<<<<<< HEAD
=======
## set_authorization(cmap_key = "5e05c500-d68d-11e9-9d3b-4f83fcec4710")
>>>>>>> 338fad1bab66c45800c700d11803e9a3a5faa782
dat = along_track("MGL1704",
                  targetTables = tabname,
                  targetVars = varname,
                  temporalTolerance = 1,
                  latTolerance = 0.25,
                  lonTolerance = 0.25,
                  depthTolerance = depthTolerance,
                  depth1 = 0,
                  depth2 = 5)
## Clean time
dat[,"time"] = as_datetime(substr(dat[,"time"], 1,20))
```

The in-situ cruise data is queried as follows (the date range of the MGL1704
Gradients 2 cruise are the relevant colocalization parameters here; everything
else is intentionally broad):

```{r query2}
tabname = "tblCruise_PAR"
varname = "par"
dat2 = get_spacetime(tableName = tabname,
                    varName = varname,
                    dt1 = '2017-05-26',
                    dt2 = '2017-06-14',
                    lat1 = -2000,
                    lat2 = 2000,
                    lon1 = -2000,
                    lon2 = 2000,
                    depth1 = 0,
                    depth2 = 100)
```

Now, join the two datasets and clean them:
```{r join}
dat = full_join(dat2, dat) %>% rename(cruise_par = par, sat_par = PAR) %>%
  mutate(time = as_datetime(time)) %>%
  arrange(time)
```

Then plot them together (after standardizing) in a single plot.

```{r plot, fig.width=10, fig.height=5}
plot(x = dat$time,
     y = dat %>% select(cruise_par) %>%scale(), type='l', ylim = c(-3,3),
     ylab = "Standardized PAR",
     xlab = "Date")
lines(x = dat$time,
      y = dat %>% select(sat_par) %>%scale(), type='l', col='red')
legend("topleft", col=c("black", "red"), lwd=2, lty = 1, legend=c("In-situ", "Remote-sensing"))
```
  
We can see that remote sensing PAR is messy (probably due to factors like cloud
cover) compared to the consistent, non-erratic daily fluctiation of in-situ PAR.
